Faithful Variable Screening for\\ High-Dimensional Convex Regression

### 2015-11-19 tex/scam/additive_faithful.tex
a
minor revisions

Because $f(x_k, \mathbf{x}_{-k})$ as a function of $x_k$ is convex, it must be that, for all $x_k \in (0,1)$ and for all $\mathbf{x}_{-k}$,
\begin{equation}
0 = f'(x_k^0, \mathbf{x}_{-k}) \leq f'(x_k, \mathbf{x}_{-k}) \leq 
    f'(x_k^1, \mathbf{x}_{-k}) = 0
\end{equation}
Therefore $f$ does not depend on $x_k$.
% Now we apply the first-order condition of convex functions to any two points $(x_k, \mathbf{x}_{-k})$ and $(x'_k, \mathbf{x}_{-k})$ where $x_k, x'_k \in [0,1]$:

% \begin{align*}
% \forall \mathbf{x}_{-k}, f(x'_k, \mathbf{x}_{-k}) & \leq f(x_k, \mathbf{x}_{-k}) 
%   + f'(x_k, \mathbf{x}_{-k}) ( x'_k - x_k) \\ 
% f(x'_k, \mathbf{x}_{-k}) &\leq f(x_k, \mathbf{x}_{-k}) \\
% \forall \mathbf{x}_{-k}, f(x_k, \mathbf{x}_{-k}) & \leq f(x'_k, \mathbf{x}_{-k}) 
%   + f'(x'_k, \mathbf{x}_{-k}) ( x_k - x'_k) \\ 
% f(x_k, \mathbf{x}_{-k}) &\leq f(x'_k, \mathbf{x}_{-k})
% \end{align*}

% We thus have that $f(x_k, \mathbf{x}_{-k}) = f(x'_k, \mathbf{x}_{-k})$ for all $\mathbf{x}_{-k}$ and all pairs $x_k, x'_k \in (0,1)$. This proves that $f$ does not depend on $x_k$.
\end{proof}

@@

\begin{remark}
In Theorem~\ref{thm:convex_faithful}, we do not assume a parametric
form for the additive components; the additive approximations may not
be faithful if we take a parametric form. For example, suppose we
approximate a mean-zero convex function $f(X)$ by a linear form $X
\beta$. The optimal linear function in the population setting is
$\beta^* = \Sigma^{-1} \trm{Cov}(X, f(X))$ where $\Sigma$ is the
covariance matrix. Suppose the $X$s are independent, have a
symmetric distribution with unit variance, and suppose $f(\mathbf{x})
= x_1^2 - \E[X_1^2]$. Then $\beta^*_1 = \E[ X_1 f(X)] = \E[ X_1^3 -
  X_1 \E[X_1^2]] = 0$.
\end{remark}

@@

Now, let $p(\mathbf{x})$ be a bounded density over $[0,1]^p$. Let
$p_\epsilon(\mathbf{x})$ be the density formed from truncating
$p(\mathbf{x})$ in $[\epsilon, 1-\epsilon]^p$. The corresponding
mixture $p_{\lambda,\epsilon}(\mathbf{x})$ then approximates
$p(\mathbf{x})$ when $\lambda$ and $\epsilon$ are both small.

@@

In the discussion so far, we have restricted our attention to
densities supported and positive on the hypercube $[0,1]^p$ to
minimize extraneous technical details. It may be possible to extend
the analysis to densities whose support is a convex and compact set so
long as the marginal density $p(x_j) > 0$ for all $x_j$ in the
support. A rigorous analysis of this however is beyond the scope of
this paper.

@@

\begin{figure*}[htp]
\vskip-10pt
	\centering
	\subfigure[Gaussian distribution]{
          \label{fig:gaussian_example}
		\centering
		{\includegraphics[width=0.4\textwidth]{figs/gaussian_example}}
	}
	\subfigure[Mixture approximation]{
          \label{fig:boundaryflat_example}
		\centering
		{\includegraphics[width=0.4\textwidth]{figs/boundaryflat_example}}
	}
\caption{Optimal additive projection of the quadratic function
  described in Example~\ref{examp:gaussian_counterexample} under both
  the Gaussian distribution described in
  Example~\ref{examp:gaussian_counterexample} and the
  approximately Gaussian mixture distribution described in
  Example~\ref{examp:boundaryflat_example}. For the mixture
  approximation, we used $b=5, \epsilon=0.3, \lambda=0.0001$ where the
  parameters are defined in
  Example~\ref{examp:boundaryflat_example}. This example shows the
  effect and the importance of the boundary flatness condition.}
\vskip-10pt
\end{figure*}

@@

Although convex functions are additively faithful---under appropriate conditions---it is difficult to
estimate the optimal additive functions $f^*_k$s as defined in
equation~\eqref{eqn:unconstrained_additive}.  The reason is that $f^*_k$ need not be
a convex function, as Example~\ref{examp:gaussian_counterexample} and Example~\ref{examp:boundaryflat_example} 
show. It may be possible to estimate $f^*_k$ via smoothing, but we prefer an approach that is free of smoothing parameters. 
%The function $f^*_k$ can be shown to be smooth but it is preferable to use
%shape-constraints whenever possible to avoid the introduction of the
%additional smoothing bandwidth.
Since the true regression function $f$ is convex, we approximate the additive model with a \emph{convex} additive model. We abuse notation and, for the rest of the paper, use the notation $f^*_k$ to represent convex additive fits:
%it is natural to ask when it is sufficient to estimate an convex additive model
\begin{equation}
\label{eqn:convex_additive}
\{ f^*_k \}_{k=1}^p = \arg\min \Big \{ 
    \E\Bigl( f(X) - \sum_{k=1}^p f_k(X_k) \Bigr)^2 \,:\, f_k \in \mathcal{C}^1, \, \E f_k(X_k) = 0 \Big \}
\end{equation}
where $\mathcal{C}^1$ is the set of univariate convex functions.

@@

\begin{lemma}
\label{cor:faithfulness_extension}
Suppose $p(\mathbf{x})$ is a density on $C=[0,1]^p$ bounded away from
$0$ and~$\infty$ satisfying the boundary flatness condition.  Let
$f(\mathbf{x})$ be a convex function with a bounded second derivative
on an open set around $C$. Let $\phi(\mathbf{x}_{-k})$ be a bounded
function that does not depend on $x_k$. Then the
unconstrained univariate function
 \begin{align}
h^*_k = \arg\min_{f_k} \mathbb{E} \Bigl[\bigl( f(X) 
           - \phi(X_{-k}) - h_k(X_k) \bigr)^2\Bigr]
\end{align}
is given by $h^*_k(x_k) = \mathbb{E}\bigl[ f(X) - \phi(X_{-k}) \,|\, x_k\bigr]$,
and $ h^*_k = 0$ implies that $\partial_{x_k} f(\mathbf{x}) = 0$.
\end{lemma}

@@

Let us suppose that $f^*_k = g^*_k = 0$. It must be then that
\begin{align*}
\argmin_{c \in \R} \E \left( f(X) - \phi(X_{-k}) - c (X_k^2 - m_k^2) \right)^2 = 0
\end{align*}
where $m_k^2 \equiv \E X_k^2$; this is because $c(x_k^2 - m_k^2)$ is either convex or concave in $x_k$ and it is centered, i.e. $\E[ X_k^2 - m_k^2] = 0$. Since the optimum has a closed form $c^* = \frac{\E\big[(f(X) -\phi(X_{-k}))(X_k^2 - m^2_k)\big]}{\E X_k^2 - m^2_k}$, we deduce that 
\begin{align*}
\E \big[ (f(X) - \phi(X_{-k})) &(X_k^2 - m_k^2) \big] \\
   &= \E[ (f(X) - \phi(X_{-k})) X^2_k] = 
 \E[ \E[ f(X) - \phi(X_{-k}) \given X_k] X^2_k] = 0
\end{align*}

@@

Because we assumed $f^* = g^* = 0$, it must be that
\[
\argmin_{c \in \R} 
\E\Big( f(X) - \phi(X_{-k}) - c\big( s_n(X_k) + \alpha(X_k^2 - m_k^2)\big) \Big)^2 = 0
\]
This is because $c\big( s_n(x_k) + \alpha(x_k^2 - m_k^2) \big)$ is
convex for $c \geq 0$ and concave for $c \leq 0$ and it is a centered
function.

Again, $c^* = \frac{\mathbb{E}[(f(X)-\phi(X_{-k}))\big( 
           s_n(X_k) + \alpha (X_k^2 - m_k^2) \big)]}{\mathbb{E}
       \big( s_n(X_k) + \alpha (X_k^2 - m_k^2) \big)^2} = 0$, so
\begin{align*}
\mathbb{E}[ (f(X)-\phi(X_{-k})) \big( s_n(X_k) + \alpha (X_k^2-m_k^2) \big) ] &= 
\mathbb{E}[ (f(X) - \phi(X_{-k})) s_n(X_k) ] \\
& = \mathbb{E}\Big[ \mathbb{E}[ f(X) - \phi(X_{-k}) \given X_k]  s_n(X_k) \Big] \\
& = \mathbb{E} h^*_k(X_k) s_n(X_k)  = 0
\end{align*}
where the first equality follows because $\E[ (f(X) - \phi(X_{-k})) (X_k^2 - m_k^2)] = 0$. \\

@@

Now we use
Lemma~\ref{cor:faithfulness_extension} with $\phi(\mathbf{x}_{-k}) =
f(\mathbf{x}) - \sum_{k' \neq k} f^*_{k'} (x_{k'})$ and conclude that
$f^*_k = 0$ and $g^*_k = 0$ together imply that $f$ does not depend on $x_k$.\\

>>>>

Because $f(x_k, \mathbf{x}_{-k})$ as a function of $x_k$ is convex, it must be that, for all $x_k \in (0,1)$ and for all $\mathbf{x}_{-k}$,
\begin{equation}
0 = f'(x_k^0, \mathbf{x}_{-k}) \leq f'(x_k, \mathbf{x}_{-k}) \leq 
    f'(x_k^1, \mathbf{x}_{-k}) = 0
\end{equation}
Therefore $f$ does not depend on $x_k$.
\end{proof}
% Now we apply the first-order condition of convex functions to any two points $(x_k, \mathbf{x}_{-k})$ and $(x'_k, \mathbf{x}_{-k})$ where $x_k, x'_k \in [0,1]$:
% \begin{align*}
% \forall \mathbf{x}_{-k}, f(x'_k, \mathbf{x}_{-k}) & \leq f(x_k, \mathbf{x}_{-k}) 
%   + f'(x_k, \mathbf{x}_{-k}) ( x'_k - x_k) \\ 
% f(x'_k, \mathbf{x}_{-k}) &\leq f(x_k, \mathbf{x}_{-k}) \\
% \forall \mathbf{x}_{-k}, f(x_k, \mathbf{x}_{-k}) & \leq f(x'_k, \mathbf{x}_{-k}) 
%   + f'(x'_k, \mathbf{x}_{-k}) ( x_k - x'_k) \\ 
% f(x_k, \mathbf{x}_{-k}) &\leq f(x'_k, \mathbf{x}_{-k})
% \end{align*}

% We thus have that $f(x_k, \mathbf{x}_{-k}) = f(x'_k, \mathbf{x}_{-k})$ for all $\mathbf{x}_{-k}$ and all pairs $x_k, x'_k \in (0,1)$. This proves that $f$ does not depend on $x_k$.

@@

\begin{remark}
In Theorem~\ref{thm:convex_faithful}, we do not assume a parametric
form for the additive components; the additive approximations may not
be faithful if we take a parametric form. For example, suppose we
approximate a mean-zero convex function $f(X)$ by a linear form $X
\beta$. The optimal linear function in the population setting is
$\beta^* = \Sigma^{-1} \trm{Cov}(X, f(X))$ where $\Sigma = X^\tran X$ is the
covariance matrix. Suppose the $X$'s are independent, centered, and have a
symmetric distribution with unit variance, and suppose $f(\mathbf{x})
= x_1^2 - \E[X_1^2]$. Then $\beta^*_1 = \E[ X_1 f(X)] = \E[ X_1^3 -
  X_1 \E[X_1^2]] = 0$.
\end{remark}

@@

Now, let $p(\mathbf{x})$ be a bounded density over $[0,1]^p$. Let
$p_\epsilon(\mathbf{x})$ be the density formed from truncating
$p(\mathbf{x})$ in $[\epsilon, 1-\epsilon]^p$ and proper re-normalization. The corresponding
mixture $p_{\lambda,\epsilon}(\mathbf{x})$ then approximates
$p(\mathbf{x})$ when $\lambda$ and $\epsilon$ are both small.

@@

In the discussion so far, we have restricted our attention to
densities supported and positive on the hypercube $[0,1]^p$ to
minimize extraneous technical details. It may also be possible to extend
the analysis to densities whose support is a convex and compact set so
long as the marginal density $p(x_j) > 0$ for all $x_j$ in the
support. A rigorous analysis of this however is beyond the scope of
this paper.

@@

\begin{figure*}[t]
\vskip-10pt
	\centering
	\subfigure[Gaussian distribution]{
          \label{fig:gaussian_example}
		\centering
		{\includegraphics[width=0.4\textwidth]{figs/gaussian_example}}
	}
	\subfigure[Mixture approximation]{
          \label{fig:boundaryflat_example}
		\centering
		{\includegraphics[width=0.4\textwidth]{figs/boundaryflat_example}}
	}
\caption{Optimal additive projection of the quadratic function
  described in Example~\ref{examp:gaussian_counterexample} under both
  the Gaussian distribution described in
  Example~\ref{examp:gaussian_counterexample} and the
  approximately Gaussian mixture distribution described in
  Example~\ref{examp:boundaryflat_example}. For the mixture
  approximation, we used $b=5, \epsilon=0.3, \lambda=0.0001$ where the
  parameters are defined in
  Example~\ref{examp:boundaryflat_example}. This example shows the
  effect and the importance of the boundary flatness condition.}
\vskip-10pt
\end{figure*}

@@

Although convex functions are additively faithful---under appropriate conditions---it is difficult to
estimate the optimal additive functions $f^*_k$s as defined in
equation~\eqref{eqn:unconstrained_additive}.  The reason is that $f^*_k$ need not be
a convex function, as Example~\ref{examp:gaussian_counterexample} and Example~\ref{examp:boundaryflat_example} 
show. It may be possible to estimate $f^*_k$ via smoothing, but we prefer an approach that is free of smoothing parameters. 
%The function $f^*_k$ can be shown to be smooth but it is preferable to use
%shape-constraints whenever possible to avoid the introduction of the
%additional smoothing bandwidth.
Since the true regression function $f$ is convex, we approximate the additive model with a \emph{convex} additive model. Without loss of generality, we assume in this section that $\E f(X) = 0$.

We abuse notation and, for the rest of the paper, use the notation $f^*_k$ to represent convex additive fits:
%it is natural to ask when it is sufficient to estimate an convex additive model
\begin{equation}
\label{eqn:convex_additive}
\{ f^*_k \}_{k=1}^p = \arg\min \Big \{ 
    \E\Bigl( f(X) - \sum_{k=1}^p f_k(X_k) \Bigr)^2 \,:\, f_k \in \mathcal{C}^1, \, \E f_k(X_k) = 0 \Big \}
\end{equation}
where $\mathcal{C}^1$ is the set of univariate convex functions.

@@

\begin{lemma}
\label{cor:faithfulness_extension}
Suppose $p(\mathbf{x})$ is a density on $C=[0,1]^p$ bounded away from
$0$ and~$\infty$ satisfying the boundary flatness condition.  Let
$f(\mathbf{x})$ be a convex function with a bounded second derivative
on an open set around $C$. Let $\phi(\mathbf{x}_{-k})$ be a bounded
function that does not depend on $x_k$. Then the
unconstrained univariate function
 \begin{align}
h^*_k = \arg\min_{h_k} \mathbb{E} \Bigl[\bigl( f(X) 
           - \phi(X_{-k}) - h_k(X_k) \bigr)^2\Bigr]
\end{align}
is given by $h^*_k(x_k) = \mathbb{E}\bigl[ f(X) - \phi(X_{-k}) \,|\, x_k\bigr]$,
and $ h^*_k = 0$ implies that $\partial_{x_k} f(\mathbf{x}) = 0$.
\end{lemma}

@@

Let us suppose that $f^*_k = g^*_k = 0$. It must be then that
\begin{align*}
\argmin_{c \in \R} \E \left( f(X) - \phi(X_{-k}) - c (X_k^2 - m_k^2) \right)^2 = 0
\end{align*}
where $m_k^2 \equiv \E X_k^2$; this is because $c(x_k^2 - m_k^2)$ is either convex or concave in $x_k$ and it is centered, i.e. $\E[ X_k^2 - m_k^2] = 0$. Since the optimum has a closed form $c^* = \frac{\E\big[(f(X) -\phi(X_{-k}))(X_k^2 - m^2_k)\big]}{(\E X_k^2 - m^2_k)^2}$, we deduce that 
\begin{align*}
\E \big[ (f(X) - \phi(X_{-k})) &(X_k^2 - m_k^2) \big] \\
   &= \E[ (f(X) - \phi(X_{-k})) X^2_k] = 
 \E[ \E[ f(X) - \phi(X_{-k}) \given X_k] X^2_k] = 0
\end{align*}

@@

Because we assumed $f^* = g^* = 0$, it must be that
\[
\argmin_{c \in \R} 
\E\Big( f(X) - \phi(X_{-k}) - c\big( s_n(X_k) - \E s_n(X_k) + \alpha(X_k^2 - m_k^2)\big) \Big)^2 = 0
\]
This is because $c\big( s_n(x_k) - \E s_n(X_k) + \alpha(x_k^2 - m_k^2) \big)$ is
convex for $c \geq 0$ and concave for $c \leq 0$ and it is a centered
function.

Again, $c^* = \frac{\mathbb{E}[(f(X)-\phi(X_{-k}))\big( 
           s_n(X_k) - \E s_n(X_k)  + \alpha (X_k^2 - m_k^2) \big)]}{\mathbb{E}
       \big( s_n(X_k) - \E s_n(X_k) + \alpha (X_k^2 - m_k^2) \big)^2} = 0$, so
\begin{align*}
\mathbb{E}[ (f(X)-\phi(X_{-k})) \big( s_n(X_k)- \E s_n(X_k) + \alpha (X_k^2-m_k^2) \big) ] &= 
\mathbb{E}[ (f(X) - \phi(X_{-k})) s_n(X_k) ] \\
& = \mathbb{E}\Big[ \mathbb{E}[ f(X) - \phi(X_{-k}) \given X_k]  s_n(X_k) \Big] \\
& = \mathbb{E} h^*_k(X_k) s_n(X_k)  = 0
\end{align*}
where the first equality follows because $\E[ (f(X) - \phi(X_{-k})) (X_k^2 - m_k^2)] = 0$. \\

@@

Now we use
Lemma~\ref{cor:faithfulness_extension} with $\phi(\mathbf{x}_{-k}) =
\sum_{k' \neq k} f^*_{k'} (x_{k'})$ and conclude that
$f^*_k = 0$ and $g^*_k = 0$ together imply that $f$ does not depend on $x_k$.\\

### 2015-11-19 tex/scam/experiments.tex
minor revisions

\subsection{Simulations}
We first illustrate our methods using a simulation of the model
\begin{equation}\nonumber
         Y_i = f_0(\x_{iS}) + \epsilon_i \quad (i=1,2,\ldots,n).
\end{equation}
Here $\x_{i}$ denotes data sample $i$ drawn from some distribution
$P$, and $f_0$ is the true regression function. The variables
$\x_{iS}$ are a subset of $\x_i$ with dimension $|S|=s$, where $S$
represents the set of relevant variables, and $\epsilon_i$ is additive
noise drawn from $\mathcal{N}(0,\sigma^2)$. For all simulations, we set
$\sigma^2$ so that the signal-to-noise ratio(SNR,
$\frac{\trm{std}(Y)}{\sigma}$) is 5. Also, for all simulations except
the sixth, we choose the set of relevant variables $S$ uniformly at
random among all variables $\{1,\ldots,p\}$.

@@

\subsubsection{Third simulation}
In the \textrm{third simulation} (Figure~\ref{Support}d,e), we use the
non-Gaussian distribution described above and set the covariance
to be $\Sigma_{ij}=\nu^{|i-j|}$ for $\nu = 0, 0.2, 0.5, 0.9$. We use the
non-additive $\Q$, same as in the first experiment, with $\alpha=0.5$
and fix $p=128, s=5$. We measure success not through exact recovery
but through faithful recovery. We say that a trial is a successful if
(1) all relevant variables were recovered and (2) fewer than $20$
variables were marked as relevant overall (true sparsity $s=5$). We
use the same $\lambda$ as before. The probabilities of success are
computed from 40 independent trials and plotted against various values
of $\nu$ in Figure~\ref{Support}(d). Additionally, for $\nu = 0.5$, we
show the number of selected variables versus the sample size as a
box plot in Figure~\ref{Support}(e). As seen, for small to
moderate correlations, AC/DC can successfully recover the relevant
variables with only a small number of false positives.

>>>>

\subsection{Simulations}
We first illustrate our methods using a simulation of the model
\begin{equation}\nonumber
         Y_i = f_0(\x_{iS}) + w_i \quad (i=1,2,\ldots,n).
\end{equation}
Here $\x_{i}$ denotes data sample $i$ drawn from some distribution
$P$, and $f_0$ is the true regression function. The variables
$\x_{iS}$ are a subset of $\x_i$ with dimension $|S|=s$, where $S$
represents the set of relevant variables, and $w_i$ is additive
noise drawn from $\mathcal{N}(0,\sigma^2)$. For all simulations, we set
$\sigma^2$ so that the signal-to-noise ratio(SNR,
$\frac{\trm{std}(Y)}{\sigma}$) is 5. Also, for all simulations except
the sixth, we choose the set of relevant variables $S$ uniformly at
random among all variables $\{1,\ldots,p\}$.

@@

\subsubsection{Third simulation}
In the \textrm{third simulation} (Figure~\ref{Support}d,e), we use the
non-Gaussian distribution described above and set the covariance
to be $\Sigma_{ij}=\nu^{|i-j|}$ for $\nu = 0, 0.2, 0.5, 0.9$. We use the
non-additive $\Q$, same as in the first experiment, with $\alpha=0.5$
and fix $p=128, s=5$. We measure success not through exact recovery
but through faithful recovery. We say that a trial is successful if
(1) all relevant variables were recovered and (2) fewer than $20$
variables were marked as relevant overall (true sparsity $s=5$). We
use the same $\lambda$ as before. The probabilities of success are
computed from 40 independent trials and plotted against various values
of $\nu$ in Figure~\ref{Support}(d). Additionally, for $\nu = 0.5$, we
show the number of selected variables versus the sample size as a
box plot in Figure~\ref{Support}(e). As seen, for small to
moderate correlations, AC/DC can successfully recover the relevant
variables with only a small number of false positives.

### 2015-11-19 tex/scam/sparsistency.tex
minor revisions

Our goal is to show variable screening consistency. That is, as
$n,p \rightarrow \infty$, $\mathbb{P}( \hat{S} = S)$ approaches 0
where $\hat{S}$ is the set of variables output by AC/DC in the
finite sample setting (Figure~\ref{fig:backfitting:algo}) and $S$ is
the set of variables output in the population
setting~\eqref{eqn:acdc_vars_pop}.

@@

Let $\mathcal{C}^1$ denote the set of univariate convex functions
supported on $[-1,1]$, 
and let  $\mathcal{C}_1^{p}$ denote the set of convex additive functions
$\mathcal{C}_1^p \equiv \{ f \,:\, f = \sum_{k=1}^p f_k, \,
   f_k \in \mathcal{C}^1 \} $.  
Let $f^*(\mathbf{x}) = \sum_{k=1}^p f^*_k(x_k)$ be the population risk
minimizer in $\mathcal{C}_1^p$, 
\begin{equation}
f^* = \arg\min_{f \in \mathcal{C}_1^p} \E\big(f_0(X) - f(X)
\big)^2.
\end{equation}
$f^*$ is the unique minimizer by Theorem~\ref{thm:acdc_faithful}. Similarly, we define $\mh \mathcal{C}^1$ as the set of univariate concave functions supported on $[-1, 1]$ and define
\begin{equation}
g^*_k = \arg\min_{g_k \in \mh \mathcal{C}^1} \E \big( f_0(X) - f^*(X)
- g_k(X_k) \big)^2.
\end{equation}
The $\hat{g}_k$s are unique minimizers as well. We let $S = \{ k =
1,\ldots,p \,:\, f^*_k \neq 0 \trm{ or } g^*_k \neq 0\}$ and let $s =
|S|$. By additive faithfulness (Theorem~\ref{thm:acdc_faithful}), it
must be that $S_0 \subset S$ and thus $s \geq s_0$. In some cases,
such as when $X_{S_0}, X_{S^c_0}$ are independent, we have $S = S_0$.
Each of our theorems will use a subset of the following assumptions:
\begin{packed_enum}
\item[A1:] $X_S, X_{S^c}$ are independent. 
\item[A2:] $f_0$ is convex with a bounded second derivative on an open set around $[-1,1]^p$. $\E f_0(X) = 0$.
\item[A3:] $\|f_0\|_\infty \leq sB$ and $\| f^*_k \|_\infty \leq B$ for all $k$.
\item[A4:] $W$ is mean-zero sub-Gaussian, independent of $X$, with scale $\sigma$; i.e., for all $t \in \R$, $\E e^{t \epsilon} \leq e^{\sigma^2 t^2 / 2}$.
\item[A5:] The density $p(\mathbf{x})$ satisfies the boundary flatness condition (Definition~\ref{defn:boundary-point}), and $0 < c_l \leq \inf p(\mathbf{x}) \leq \sup p(\mathbf{x}) \leq c_u < \infty$ for two constants $c_l, c_u$. 
\end{packed_enum}
By assumption A1, $f^*_k$ must be zero for $k\notin S$.
We define $\alpha_{+}, \alpha_{-}$ as a measure of the signal strength of the weakest variable:
\begin{align}
\alpha_{+} &= \min_{f \in \mathcal{C}_1^p \,:\, \textrm{supp}(f) \subsetneq \textrm{supp}(f^*)} 
       \Big\{ \mathbb{E} \big( f_0(X) - f(X) \big)^2 - 
        \mathbb{E} \big( f_0(X) - f^*(X) \big)^2  \Big\} \label{eqn:signal_level_defn} \\
\alpha_{-} &=   \min_{k \in S \,:\, g^*_k \neq 0}
      \Big\{ \mathbb{E} \big( f_0(X) - f^*(X) \big)^2 - 
    \mathbb{E} \big( f_0(X) - f^*(X) - g^*_k(X_k) \big)^2 \Big\} \nonumber
\end{align}
The term $\alpha_+$ is a lower bound on the excess risk incurred by
any additive convex function whose support is strictly smaller than
$f^*$. $\alpha_+$ is achieved by some $f \neq f^*$ because the set
$\{f\in \mathcal{C}_1^p\,:\, \trm{supp}(f) \subsetneq
\textrm{supp}(f^*)\}$ is a finite union of closed convex
sets. $\alpha_+ > 0$ since $f^*$ is the unique risk
minimizer. Likewise, $\alpha_-$ lower bounds the excess risk of any
decoupled concave fit of the residual $f_0 - f^*$ that is strictly
more sparse than the optimal decoupled concave fit $\{\hat{g}_k^*\}$;
$\alpha_- > 0$ by the uniqueness of $\{g^*_k\}$ as well. These
quantities play the role of the absolute value of the smallest nonzero
coefficient in the true linear model in lasso theory.  Intuitively, if
$\alpha_{+}$ is small, then it is easier to make a false omission in
the additive convex stage of the procedure. If $\alpha_{-}$ is small,
then it is easier to make a false omission in the decoupled concave
stage of the procedure. If $X$ is independent, then $\alpha_+$ can be
simplified to $\min_{k \,:\, f^*_k \neq 0} \E f^*_k(X)^2$ and
$\alpha_-$ becomes unnecessary.

@@

\begin{theorem}[Controlling false positives]
\label{thm:false_positive}
Suppose assumptions A1-A5 hold. Define $\tilde{\sigma} \equiv \max(\sigma, B)$ and define $\mathsf{range}_k = X_{\pi_k(n)k} - X_{\pi_k(1)k}$. Suppose $p \leq O\big( \exp(c n) \big)$ and $n \geq C$ for some positive constants $C$ and $0 < c < \frac{c_l}{32}$.  Suppose also
\begin{equation}
\lambda_n \geq 768 s \tilde{\sigma}  \sqrt{ \frac{\log^2 np}{n}}.
\end{equation}  
Then with probability at least $ 1 - \frac{24}{n}$, for all $k \in
S^c$, for all $i=1,\ldots,n$,
\begin{equation}
\lambda_n \geq \mathsf{range}_k \Big| \frac{32}{n}\hat{r}^\tran \mathbf{1}_{(i':n)_k} \Big| ,
\end{equation}
$\max_{i'} \frac{X_{\pi_k(i'+1)k} - X_{\pi_k(i')k}}{\mathsf{range}_k} \leq \frac{1}{16}$, $\mathsf{range}_k \geq 1$, and both the AC solution $\hat{f}_k$ from optimization~\eqref{opt:alternate_opt} and the DC solution $\hat{g}_k$ from optimization~\eqref{opt:alternate_opt_concave} are zero. 
\end{theorem}

>>>>

Our goal is to show variable screening consistency. That is, as
$n,p \rightarrow \infty$, $\mathbb{P}( \hat{S} = S)$ approaches 1
where $\hat{S}$ is the set of variables output by AC/DC in the
finite sample setting (Figure~\ref{fig:backfitting:algo}) and $S$ is
the set of variables output in the population
setting~\eqref{eqn:acdc_vars_pop}.

@@

Let $\mathcal{C}^1$ denote the set of univariate convex functions
supported on $[-1,1]$, 
and let  $\mathcal{C}_1^{p}$ denote the set of convex additive functions
$\mathcal{C}_1^p \equiv \{ f \,:\, f = \sum_{k=1}^p f_k, \,
   f_k \in \mathcal{C}^1 \} $.  
Let $f^*(\mathbf{x}) = \sum_{k=1}^p f^*_k(x_k)$ be the population risk
minimizer in $\mathcal{C}_1^p$, 
\begin{equation}
f^* = \arg\min_{f \in \mathcal{C}_1^p} \E\big(f_0(X) - f(X)
\big)^2.
\end{equation}
$f^*$ is the unique minimizer by Theorem~\ref{thm:acdc_faithful}. Similarly, we define $\mh \mathcal{C}^1$ as the set of univariate concave functions supported on $[-1, 1]$ and define
\begin{equation}
g^*_k = \arg\min_{g_k \in \mh \mathcal{C}^1} \E \big( f_0(X) - f^*(X)
- g_k(X_k) \big)^2.
\end{equation}
The $\hat{g}_k$s are unique minimizers as well. We let $S = \{ k =
1,\ldots,p \,:\, f^*_k \neq 0 \trm{ or } g^*_k \neq 0\}$ and let $s =
|S|$. By additive faithfulness (Theorem~\ref{thm:acdc_faithful}), it
must be that $S_0 \subset S$ and thus $s \geq s_0$. In some cases,
such as when $X_{S_0}, X_{S^c_0}$ are independent, we have $S = S_0$.
Each of our theorems will use a subset of the following assumptions:
\begin{packed_enum}
\item[A1:] $X_S, X_{S^c}$ are independent. 
\item[A2:] $f_0$ is convex with a bounded second derivative on an open set around $[-1,1]^p$. $\E f_0(X) = 0$.
\item[A3:] $\|f_0\|_\infty \leq sB$ and $\| f^*_k \|_\infty \leq B$ for all $k$.
\item[A4:] $W$ is mean-zero sub-Gaussian, independent of $X$, with scale $\sigma$; i.e., for all $t \in \R$, $\E e^{t W} \leq e^{\sigma^2 t^2 / 2}$.
\item[A5:] The density $p(\mathbf{x})$ satisfies the boundary flatness condition (Definition~\ref{defn:boundary-point}), and $0 < c_l \leq \inf p(\mathbf{x}) \leq \sup p(\mathbf{x}) \leq c_u < \infty$ for two constants $c_l, c_u$. 
\end{packed_enum}
By assumption A1, $f^*_k$ must be zero for $k\notin S$.
We define $\alpha_{+}, \alpha_{-}$ as a measure of the signal strength of the weakest variable:
\begin{align}
\alpha_{+} &= \min_{f \in \mathcal{C}_1^p \,:\, \textrm{supp}(f) \subsetneq \textrm{supp}(f^*)} 
       \Big\{ \mathbb{E} \big( f_0(X) - f(X) \big)^2 - 
        \mathbb{E} \big( f_0(X) - f^*(X) \big)^2  \Big\} \label{eqn:signal_level_defn} \\
\alpha_{-} &=   \min_{k \in S \,:\, g^*_k \neq 0}
      \Big\{ \mathbb{E} \big( f_0(X) - f^*(X) \big)^2 - 
    \mathbb{E} \big( f_0(X) - f^*(X) - g^*_k(X_k) \big)^2 \Big\} \nonumber
\end{align}
The term $\alpha_+$ is a lower bound on the excess risk incurred by
any additive convex function whose support is strictly smaller than
$f^*$. $\alpha_+$ is achieved by some $f \neq f^*$ because the set
$\{f\in \mathcal{C}_1^p\,:\, \trm{supp}(f) \subsetneq
\textrm{supp}(f^*)\}$ is a finite union of closed convex
sets. $\alpha_+ > 0$ since $f^*$ is the unique risk
minimizer. Likewise, $\alpha_-$ lower bounds the excess risk of any
decoupled concave fit of the residual $f_0 - f^*$ that is strictly
more sparse than the optimal decoupled concave fit $\{\hat{g}_k^*\}$;
$\alpha_- > 0$ by the uniqueness of $\{g^*_k\}$ as well. These
quantities play the role of the absolute value of the smallest nonzero
coefficient in the true linear model in lasso theory.  Intuitively, if
$\alpha_{+}$ is small, then it is easier to make a false omission in
the additive convex stage of the procedure. If $\alpha_{-}$ is small,
then it is easier to make a false omission in the decoupled concave
stage of the procedure. If $X$ is independent, then $\alpha_+$ can be
simplified to $\min_{k \,:\, f^*_k \neq 0} \E f^*_k(X)^2$ and
$\alpha_-$ becomes unnecessary (see section~\ref{sec:prod_density_simplification} in the supplementary material).

@@

\begin{theorem}[Controlling false positives]
\label{thm:false_positive}
Suppose assumptions A1-A5 hold. Define $\tilde{\sigma} \equiv \max(\sigma, B)$ and define $\mathsf{range}_k = X_{\pi_k(n)k} - X_{\pi_k(1)k}$. Suppose $p \leq O\big( \exp(c n) \big)$ and $n \geq C$ for some positive constants $C$ and $0 < c < \frac{c_l}{32}$.  Suppose also
\begin{equation}
\lambda_n \geq 768 s \tilde{\sigma}  \sqrt{ \frac{\log^2 np}{n}}.
\end{equation}  
Then with probability at least $ 1 - \frac{24}{n}$, for all $k \in
S^c$, for all $i'=1,\ldots,n$,
\begin{equation}
\lambda_n \geq \mathsf{range}_k \Big| \frac{32}{n}\hat{r}^\tran \mathbf{1}_{(i':n)_k} \Big| ,
\end{equation}
$\max_{i'} \frac{X_{\pi_k(i'+1)k} - X_{\pi_k(i')k}}{\mathsf{range}_k} \leq \frac{1}{16}$, $\mathsf{range}_k \geq 1$, and both the AC solution $\hat{f}_k$ from optimization~\eqref{opt:alternate_opt} and the DC solution $\hat{g}_k$ from optimization~\eqref{opt:alternate_opt_concave} are zero. 
\end{theorem}
Here, we use $\mathbf{1}_{(i':n)}$ to denote a vector that is 1 on the $i'$-th to the $n$-th coordinates and 0 elsewhere.
